---
output:
  pdf_document: default
  html_document: default
---
# IE - Master in Business Analytics and Big Data
Data Visulization

Prof: Jaime De Mora Navarro 

Group H: Amal AlMutawa | Alan Carton | Ryan Chiu | Manuel Ehmann | Sabrina Mirtcheva | Harpeet Singh Perhar | Andrea Savlati


# Initialization

Import libraries

```{r}
# Load libaries for graphs & plots
library(ggplot2)
library(dbplyr)
library(data.table)
library(tidyverse)
library(scales)
library(ggpubr)
```

Load the data

```{r} 
#change path to match yours
transactions = read.csv("//Users/amal/Desktop/Group\ H\ DV\ /madrid_transactions.csv", header=T, na.strings=c("","NA"))

```

Explore the dataset.

```{r}
summary(transactions)
str(transactions)
```

## Produce Graphs

# slide 3: What is the most active category - in terms of transactions? 
NOTE: change value of heat_map to produce different graphs & change title value

```{r }
# Create filter for transaction amount
# Value to change is here
heat_map <- dplyr::filter(transactions, amount < 51)

# Create heatmap, defining range
longdf = heat_map %>%
     group_by(category) %>%
     complete(hour = 1:23 )

# Create plot, add axis labels and range, define font colors and aesthetics 
# Change title for amount of filter
active_cat_plot <- ggplot(longdf , aes(hour, category, fill = amount)) +
  geom_tile() + labs(title ="Spending Throughout the Day for Transactions < EUR 50", 
                     x = "Hour of the Day", y = "Category") +
  scale_fill_continuous(name = "Amount Spent") +
  scale_y_discrete(expand=c(0,0)) + scale_x_discrete(limits = c(1:23), expand=c(0,0)) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(family="Gill Sans Light", face="bold", color="gray28", size=8), 
        axis.text.y = element_text(family="Gill Sans Light", face="bold", color="gray28", size=8)) +
  theme(plot.title = element_text(family="Gill Sans Light", color="gray28", size=14, face="bold"),
        axis.title.x = element_text(family="Gill Sans Light", color="gray28", size=10, face="bold"),
        axis.title.y = element_text(family="Gill Sans Light", color="gray28", size=10, face="bold")) +
    scale_fill_continuous(na.value = "grey77", high = "#132B43", low = "#56B1F7") 

active_cat_plot

```

# Slide 4: Is there a particular time in which tourists are buying?

```{r }

# Add a column indicating which category should be highlighted in red
transactions <- transactions %>% mutate(ToHighlight = ifelse(hour == 17, "yes", "no" ) )

# Create plot, add title, axes labels, assign histogram color, define x-axis scale
My_plot <- ggplot(transactions,aes(x=hour, y=amount, fill = ToHighlight )) +
  geom_bar(stat="sum", na.rm=TRUE) + ggtitle("Amount Spent Per Hour") + 
  xlab("Hours of the day") + ylab("Total Amount Spent in EUR") + 
  scale_fill_manual(values = c( "yes"="red", "no"="steelblue" ), guide = FALSE ) + 
  scale_y_continuous(label=comma, expand = c(0,0),limits = c(0,80000)) + 
  scale_x_continuous(breaks=seq(0,24,2)) + 
  theme(panel.background = element_blank(), axis.line = element_line(colour = "black"))

# Add Title, Axis Labels & Colors, remove legend
My_plot <- My_plot + theme(plot.title = element_text(family="Gill Sans Light", 
                                                     color="gray28", size=16, face="bold"),
                           axis.title.x = element_text(family="Gill Sans Light", 
                                                       color="gray28", size=10, face="bold"),
                           axis.title.y = element_text(family="Gill Sans Light", 
                                                       color="gray28", size=10, face="bold")) +
  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none")

My_plot

```

# Slide 5:  What is the most active category - in terms of transactions? 

```{r }

# Create data table, and find count of number of transactions
category_n <- as.data.table(transactions)
category_n <- category_n[, .(count= .N, var = sum(amount)), by ="category" ]
category_n <- category_n[order(count, decreasing = TRUE)]

# Filter top 7 categories
category_top7 <- category_n[0:7]
  
# Create plot
top7_plot <- ggplot(category_top7, aes(x = category, y = count))

# Create scatterplot, add title and y-axis label
top7_plot <- top7_plot + geom_jitter(aes(col=category, size=count)) +
    ggtitle("Top 7 Categories by Number of Transactions") + ylab("Number of Transactions") + 
  theme(panel.background = element_blank(), axis.line = element_line(colour = "black"))
  
# Set scale range, remove x-axis, set y-scale range, add commas to y-axis, define title aestetics
top7_plot <- top7_plot + scale_size(range=c(12,30), breaks=c(45,60)) +
  theme(axis.title.x=element_blank(), 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + 
     scale_y_continuous(limits = c(0,6500), label=comma) +
     theme(plot.title = element_text(family="Gill Sans Light", 
                                                     color="gray28", size=16, face="bold", hjust=2),
           axis.title.y = element_text(family="Gill Sans Light", 
                                       color="gray28", size=10, face="bold")) 

top7_plot

```

# Slide 6: Is there any relationship between day of the week and consumption?

```{r}

# Create data table, and find mean
day_mean <- as.data.table(transactions)
day_mean <- day_mean[, mean(amount), by ="weekday" ]

# Create plot, and add title
Thurs_Fri_plot <- ggplot(day_mean, aes(x=reorder(weekday, desc(weekday)), y=V1)) +
  geom_bar(stat="identity", aes(fill = weekday)) + 
  ggtitle("Spending on Thursday vs Friday") + 
  scale_fill_manual(values = c("slategray3", "coral1")) + 
  theme(panel.background = element_blank(), axis.line = element_line(colour = "black"))

# Change labels, color of axes; remove x-axis, add white text on top of bars
Thurs_Fri_plot <- Thurs_Fri_plot + 
  ylab("Average Total Spending") + 
  theme(plot.title = element_text(family="Gill Sans Light", color="gray28", size=16, face="bold"),
        axis.title.x = element_text(color="gray28", size=10, face="bold"),
        axis.title.y = element_text(family="Gill Sans Light", color="gray28", size=12, face="bold")) +
  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  geom_text(label=c("Thursday", "Friday"), vjust=9, 
            color = "white", position = position_dodge(1), size=8,
            aes(y = V1, label = V1)) + 
   scale_y_continuous(expand = c(0,0), limits = c(0,120))

#geom_text(aes(y = Value + 2 * sign(Value), label = Value), 
  #            position = position_dodge(width = 0.9), 
  #            size = 3.5 , angle = 90)

Thurs_Fri_plot
```

# Slide 6: Is there any relationship between day of the week and consumption - by category? 

```{r }

# Create data table, and find mean of all transactions
day_spending <- as.data.table(transactions)
day_spending <- day_spending[, mean(amount), by = c("weekday", "category")]

# Reorder y-lab
transactions$category <- factor(transactions$category, levels = rev(levels(transactions$category)))
levels(transactions$category)

# Reorder legend
transactions$weekday <- factor(transactions$weekday, levels = rev(levels(transactions$weekday)))
levels(transactions$weekday)

# Plot the graph, flip axis
THF_plot <- ggplot(day_spending, aes(x = category, y = V1, fill = weekday, label=weekday)) + 
  geom_histogram(stat = "identity", aes(col=weekday),  
                 position = position_dodge(width = 0.6), 
                 width=0.5, size=0) +
  xlab("Category") + ylab("Average Spending") + 
  scale_y_continuous(label=comma, expand = c(0,0), limits = c(0,500)) +
  coord_flip() + ggtitle("Spending on Thursday vs Friday, by Category") + 
  theme(panel.background = element_blank(), axis.line = element_line(colour = "black"))
  
# Change labels, color of axes and title, add legend spacing
THF_plot <- THF_plot + scale_fill_manual(values=c("coral1", "slategray3"), 
                                         labels = c("Thursday", "Friday"))+
  theme(legend.title = element_blank()) + 
  theme(plot.title = element_text(family="Gill Sans Light", 
                                  color="gray28", size=14, face="bold"),
         axis.title.x = element_text(family="Gill Sans Light", 
                                     color="gray28", size=10, face="bold"),
         axis.title.y = element_text(family="Gill Sans Light", 
                                     color="gray28", size=10, face="bold")) + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x=element_text(angle=0, hjust=1), legend.spacing.x = unit(0.2, 'cm')) + 
  scale_colour_discrete(guide=FALSE)

THF_plot
```

# Slide 7: Which nationality has the highest average ticket?

```{r }

# Create data table, and find mean of number of transactions, show top 20
countries_mean <- as.data.table(transactions)
countries_mean <- countries_mean[, mean(amount), by ="customer_country" ]
countries_mean <- countries_mean[order(V1, decreasing = TRUE)]
top_20_countries <- countries_mean[0:20]

# Create plot, define axes labels, define color of geom_point
top_nationality_plot <- ggplot(top_20_countries, aes(x = reorder(customer_country, -V1), y = V1)) +
  geom_point(size = 3, color = "dodgerblue3") + 
  xlab("Country") + ylab("Average Spending (EUR)") + theme_classic() +
  geom_segment(aes(x=customer_country,xend= customer_country, y=0, yend = V1)) + 
  ggtitle("Top 20 Countries With Highest Spending") + 
  scale_y_continuous(label=comma, expand = c(0,0),limits = c(0,550)) +  
  theme(plot.title = element_text(family="Gill Sans Light", 
                                                        color="gray28", size=16, face="bold"),
                              axis.title.x = element_text(family="Gill Sans Light",
                                                          color="gray28", size=10, face="bold"),
                              axis.title.y = element_text(family="Gill Sans Light", 
                                                          color="gray28", size=10, face="bold")) + 
  theme(plot.title = element_text(hjust = 0.5))

top_nationality_plot 

```

# Slide 8: How much are EU countries spending in total?  

```{r }

# Create list of EU countries
EU_codes <- c("AT", "BE", "BG", "HR", "CY", "CZ", "DK", "EE", "FI", "FR", "GR", "HU", "IE", "IT", "LV", "LT", "LU", "MT", "NL", "PL", "PT", "RO", "SK", "SI", "SE")

# Create dataframe with all the EU countries
EU_countries <- transactions
EU_countries <- dplyr::filter(EU_countries, customer_country == EU_codes)

# Create datatable, and calculate the mean of each EU country
EU_countries_mean <- as.data.table(EU_countries)
EU_countries_mean <- EU_countries_mean[, mean(amount), by ="customer_country" ]
EU_mean <- mean(EU_countries$amount)
  
# Create a new column for above/below average
EU_countries_mean$c_type <- ifelse(EU_countries_mean$V1 < EU_mean, "below", "above") 

# Create plot, change aesthetics of axes and title, center title, define legend name
ggplot(EU_countries_mean, aes(x= reorder(customer_country, V1), y=V1, label=customer_country)) + 
  geom_bar(stat='identity', aes(fill=c_type), width=.5)  + 
  xlab("Country") + ylab("Average Spending (EUR)") + 
  scale_fill_manual(name="EU Spending", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="#00ba38", "below"="#f8766d")) + 
  labs(title= "European Countries - Average Spending Habits") +
  coord_flip() + theme_classic() + scale_y_continuous(expand=c(0,0), label=comma) + 
  theme(plot.title = element_text(family="Gill Sans Light", 
                                  color="gray28", size=14, face="bold", hjust = 0.5),
  axis.title.x = element_text(family="Gill Sans Light", 
                              color="gray28", size=10, face="bold"),
  axis.title.y = element_text(family="Gill Sans Light", 
                              color="gray28", size=10, face="bold"))

```
  
# Slide 8: How much are EU countries spending on fashion?  

```{r }

# Create new dateframe filtered by EU countries for only Fashion
EU_countries_Fashion <- transactions
EU_countries_Fashion <- dplyr::filter(EU_countries_Fashion, category == "Fashion & Shoes")
EU_countries_Fashion <- dplyr::filter(EU_countries_Fashion, customer_country == EU_codes)

# Calculate the mean of each EU country spending in fashion
EU_countries_mean <- as.data.table(EU_countries_Fashion)
EU_countries_mean <- EU_countries_mean[, sum(amount), by ="customer_country" ]

# Fashion Mean of EU countries
EU_mean <- mean(EU_countries_Fashion$amount)
  
# Create a new column for above/below average
EU_countries_mean$c_type <- ifelse(EU_countries_mean$V1 < EU_mean, "below", "above") 

# Create plot, change aesthetics of axes and title, center title, define legend name
ggplot(EU_countries_mean, aes(x= reorder(customer_country, V1), y=V1, label=customer_country)) + 
  geom_bar(stat='identity', aes(fill=c_type), width=.5)  + xlab("Country") + ylab("Average Spending (EUR)") + 
  scale_fill_manual(name="EU Spending", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="#00ba38", "below"="#f8766d")) + 
  labs(title= "European Countries - Spening on Fashion") +  coord_flip() + theme_classic() +
  scale_y_continuous(expand=c(0,0), label=comma) +
  theme(plot.title = element_text(family="Gill Sans Light", 
                                  color="gray28", size=14, face="bold", hjust = 0.5),
  axis.title.x = element_text(family="Gill Sans Light", 
                              color="gray28", size=10, face="bold"),
  axis.title.y = element_text(family="Gill Sans Light", 
                              color="gray28", size=10, face="bold"))
```   